---
title: "Homework 4 Assignment"
author: "Charlie Marcou, Carrie Mecca, Jasmine Zhang, and Jessie Bustin"
fontsize: 10 pt
output: 
    pdf_document:
        fig_width: 6
        fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      #include = TRUE, 
                      fig.width = 6, fig.height = 4,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE,
                      digits = 3,
                      width = 48) 
```


We’ll explore casts for ‘drama’ movies from 1980-1999. I’ve limited the data to actors in more than ten productions over this time period (and to movies with more than ten actors).

Each question is 10 points. **5 questions** in total

## Starter code

See actors example code and data.
```{r}
## actors network example
library(igraph)

### GRAPH
## read in a graph in the `graphml' format: xml for graphs.
## it warns about pre-specified ids, but we want this here
## (these ids match up with the castlists in movies.txt)

actnet <- read.graph("actors.graphml",format="graphml")
### TRANSACTION
## read in the table of actor ids for movies
## this is a bit complex, because the movie names
## contain all sorts of special characters.

movies <- read.table("movies.txt", sep="\t", 
	row.names=1, as.is=TRUE, comment.char="", quote="")
## it's a 1 column matrix.  treat it like a vector
movies <- drop(as.matrix(movies))

## each element is a comma-separated set of actor ids.  
## use `strsplit' to break these out
movies <- strsplit(movies,",")

## and finally, match ids to names from actnet
casts <- lapply(movies, 
	function(m) V(actnet)$name[match(m,V(actnet)$id)])

## check it
casts['True Romance']

## format as arules transaction baskets
library(arules)
casttrans <- as(casts, "transactions")

## Set up STM information
castsize <- unlist(lapply(casts, function(m) length(m)))

## see ?rep.int: we're just repeating movie names for each cast member
acti <- factor(rep.int(names(casts),times=castsize))

## actors
actj <- factor(unlist(casts), levels=V(actnet)$name)

## format as STM (if you specify without `x', its binary 0/1)
actmat <- sparseMatrix(i=as.numeric(acti),j=as.numeric(actj),
		dimnames=list(movie=levels(acti),actor=levels(actj)))

## count the number of appearences by actor
nroles <- colSums(actmat)
names(nroles) <- colnames(actmat)
```

## Question 1
 The actors network has an edge if the two actors were in the same movie. Plot the entire actors network.
 
```{r}
plot(actnet, vertex.frame.color=0, vertex.size=3, edge.arrow.size=.2)
```

## Question 2
Plot the neighborhoods for “Bacon, Kevin” at orders 1-3. How does the size of the network change with order?

The network gets larger as order increases.
```{r}
#one link away
edgemat <- cbind(get.edgelist(actnet)[,1], get.edgelist(actnet)[,2])
actlink <- graph.edgelist(edgemat)
bacon <- graph.neighborhood(actlink, 1, V(actlink)["Bacon, Kevin"],mode="in")[[1]]
plot(bacon,  vertex.frame.color=0, edge.arrow.width=.3)
gsize(bacon) #edge count

#two links away
bacon2 <- graph.neighborhood(actlink, 2, V(actlink)["Bacon, Kevin"],mode="in")[[1]]
plot(bacon2,  vertex.frame.color=0, edge.arrow.width=.3)
gsize(bacon2)

#three links away
bacon3 <- graph.neighborhood(actlink, 3, V(actlink)["Bacon, Kevin"],mode="in")[[1]]
plot(bacon3,  vertex.frame.color=0, edge.arrow.width=.3)
gsize(bacon3)
```

## Question 3
Who were the most common actors? Who were most connected? Pick a pair of actors and describe the shortest path between them.

We can see that the most common actors in the dataset are "Zivojinovic, Velimir 'Bata'", "Jeremy, Ron", "Doll, Dora", "Dobtcheff, Vernon",  and "Berléand, François". 

##Describe shortest path between tom hanks/kevin bacon.
```{r}
head(nroles[order(-nroles)], 5) #largest nroles

distances(actnet, v=V(actnet)["Bacon, Kevin"], to=V(actnet)["Hanks, Tom"], weights=NA)#started shortest path

#copied some of her code from her firenze example
# All shortest paths
all_shortest_paths(actnet, from="Bacon, Kevin", to="Hanks, Tom")
# pick one
BaconHanksPaths <- get.shortest.paths(actnet, from="Bacon, Kevin", to="Hanks, Tom")
V(actnet)$name[BaconHanksPaths$vpath[[1]]]
#not sure what to describe here bc the edge length is one, and it's not like we can describe what movies they shared based on this. I guess we could pick a less common pair that hopefully have two edges to describe better.

#visualizing, but it's too dense to see
E(actnet)$width <- 2
E(actnet)$color <- "grey"
E(actnet, path=BaconHanksPaths$vpath[[1]])$color <- "green"
plot(actnet)
```

## Question 4
Find pairwise actor-cast association rules with at least 0.01% support and 10% confidence. Describe what you find.

We find 92,555 rules created. We can see that if we plot betweenness and degree, that these measures of connectivity seem to have a positive association with each other. ##idk how relevant that is to say. Dobtcheff, Vernon has the highest betweeness of all actors. Which indicates that he his frequently the shortest path between actors when using these actor-cast association rules. This high centrality makes sense considering that Dobtcheff was shown to be one of the most frequent actos in the dataset in Q3.  Meanwhile, Corrêa e Castro, Cláudio has the highest degree, which indicates he has the most rules relating to other actors. 

```{r}
actorrules <- apriori(casttrans, parameter=list(support=.0001, confidence=.1, maxlen=2))
inspect(actorrules)

pairs <- labels(actorrules)
pairs <- gsub("\\{|\\}","",pairs)
pairs <- strsplit(pairs," => ")
pairs <- do.call(rbind,pairs)
pairs <- pairs[pairs[,1]!="",]

rulenet <- graph.edgelist(pairs)
rulenet <- as.undirected(rulenet)

plot(rulenet, vertex.label=NA, vertex.size=3, edge.curved=FALSE)


## look at some measures of connectivity
mbetween <- betweenness(rulenet)
mdegree <- degree(rulenet)
which.max(mbetween)
which.max(mdegree)

## and raw popularity
#playcount <- table(lastfm$artist)[names(mdegree)]

plot(mdegree,mbetween,log="xy")
#plot(as.numeric(playcount),mdegree,log="xy")

##stuff with playcount, would be like actcount for us, but I don't think our stuff is in the same format as in her lastfm.R example. not sure we want to lose time trying to figure that out. 
```

## Question 5
Let's zoom in at "Sheen, Martin". Build a regression model to predict his presence in a production from other actors' presences. Who are the top 5 actors that increase Martin Sheen's chance in a production? Compare your finding with the one from association rules? 

After fitting the LASSO regression model to predict Martin Sheen's presence in a movie, we find that the top 5 actors to increase his odds of participation are: Abraham Murray, Rick Warner, Sydney Penny, Marlee Martin and Matt Clark.

We can compare this to the rules of association that we just created by graphing Martin Sheen's neighbors in the network of rules. We see that Abraham Murray, Rick Warner, Sydney Penny, Marlee Martin, and Matt Clark are all neighbors. Martin Sheen has more than just five neighbors that he is associated with who likely also are important predictors in the regression model.

```{r}
target <- as.matrix(as.integer(actmat[,"Sheen, Martin"]))
predictors <- actmat[,!(colnames(actmat) %in% "Sheen, Martin")]

set.seed(100)
library(glmnet)
logit.fit <- cv.glmnet(predictors,target, family="binomial")
logit_coef <- coef(logit.fit, "lambda.min")
nonzero<- which(logit_coef!=0,arr.ind=FALSE)
nonzero_coeff<- as.matrix(logit_coef[nonzero,])
head(nonzero_coeff[order(-nonzero_coeff),], 5)

##comparing to Sheen Martin Association Rules by graphing neighbors
nei <- graph.neighborhood(rulenet, 1, V(rulenet)["Sheen, Martin"])[[1]]
plot(nei, edge.curved=FALSE)


```

