---
title: "Homework 4 Assignment"
author: "Charlie Marcou, Carrie Mecca, Jasmine Zhang, and Jessie Bustin"
fontsize: 10 pt
output: 
    pdf_document:
        fig_width: 6
        fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      #include = TRUE, 
                      fig.width = 6, fig.height = 4,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE,
                      digits = 3,
                      width = 48) 
```


We’ll explore casts for ‘drama’ movies from 1980-1999. I’ve limited the data to actors in more than ten productions over this time period (and to movies with more than ten actors).

Each question is 10 points. **5 questions** in total

## Starter code

See actors example code and data.
```{r}
## actors network example
library(igraph)

### GRAPH
## read in a graph in the `graphml' format: xml for graphs.
## it warns about pre-specified ids, but we want this here
## (these ids match up with the castlists in movies.txt)

actnet <- read.graph("actors.graphml",format="graphml")
### TRANSACTION
## read in the table of actor ids for movies
## this is a bit complex, because the movie names
## contain all sorts of special characters.

movies <- read.table("movies.txt", sep="\t", 
	row.names=1, as.is=TRUE, comment.char="", quote="")
## it's a 1 column matrix.  treat it like a vector
movies <- drop(as.matrix(movies))

## each element is a comma-separated set of actor ids.  
## use `strsplit' to break these out
movies <- strsplit(movies,",")

## and finally, match ids to names from actnet
casts <- lapply(movies, 
	function(m) V(actnet)$name[match(m,V(actnet)$id)])

## check it
casts['True Romance']

## format as arules transaction baskets
library(arules)
casttrans <- as(casts, "transactions")

## Set up STM information
castsize <- unlist(lapply(casts, function(m) length(m)))

## see ?rep.int: we're just repeating movie names for each cast member
acti <- factor(rep.int(names(casts),times=castsize))

## actors
actj <- factor(unlist(casts), levels=V(actnet)$name)

## format as STM (if you specify without `x', its binary 0/1)
actmat <- sparseMatrix(i=as.numeric(acti),j=as.numeric(actj),
		dimnames=list(movie=levels(acti),actor=levels(actj)))

## count the number of appearences by actor
nroles <- colSums(actmat)
names(nroles) <- colnames(actmat)
```

## Question 1
 The actors network has an edge if the two actors were in the same movie. Plot the entire actors network.
 
```{r}
plot(actnet, vertex.frame.color=0, vertex.size=3, edge.arrow.size=.2)
```

## Question 2
Plot the neighborhoods for “Bacon, Kevin” at orders 1-3. How does the size of the network change with order?

The network gets larger as order increases.
```{r}
#one link away
edgemat <- cbind(get.edgelist(actnet)[,1], get.edgelist(actnet)[,2])
actlink <- graph.edgelist(edgemat)
bacon <- graph.neighborhood(actlink, 1, V(actlink)["Bacon, Kevin"],mode="in")[[1]]
plot(bacon,  vertex.frame.color=0, edge.arrow.width=.3)
gsize(bacon) #edge count

#two links away
bacon2 <- graph.neighborhood(actlink, 2, V(actlink)["Bacon, Kevin"],mode="in")[[1]]
plot(bacon2,  vertex.frame.color=0, edge.arrow.width=.3)
gsize(bacon2)

#three links away
bacon3 <- graph.neighborhood(actlink, 3, V(actlink)["Bacon, Kevin"],mode="in")[[1]]
plot(bacon3,  vertex.frame.color=0, edge.arrow.width=.3)
gsize(bacon3)
```

## Question 3
Who were the most common actors? Who were most connected? Pick a pair of actors and describe the shortest path between them.

```{r}
head(nroles[order(-nroles)], 5) #largest nroles

distances(actnet, v=V(actnet)["Bacon, Kevin"], to=V(actnet)["Hanks, Tom"], weights=NA)#started shortest path

#copied some of her code from her firenze example
# All shortest paths
all_shortest_paths(actnet, from="Bacon, Kevin", to="Hanks, Tom")
# pick one
BaconHanksPaths <- get.shortest.paths(actnet, from="Bacon, Kevin", to="Hanks, Tom")
V(actnet)$name[BaconHanksPaths$vpath[[1]]]
#not sure what to describe here bc the edge length is one, and it's not like we can describe what movies they shared based on this. I guess we could pick a less common pair that hopefully have two edges to describe better.

#visualizing, but it's too dense to see
E(actnet)$width <- 2
E(actnet)$color <- "grey"
E(actnet, path=BaconHanksPaths$vpath[[1]])$color <- "green"
plot(actnet)
```

## Question 4
Find pairwise actor-cast association rules with at least 0.01% support and 10% confidence. Describe what you find.
```{r}
##not in right format, will need to fix
actorrules <- apriori(actnet, 
	parameter=list(support=.01, confidence=.5, maxlen=3))
```

## Question 5
Let's zoom in at "Sheen, Martin". Build a regression model to predict his presence in a production from other actors' presences. Who are the top 5 actors that increase Martin Sheen's chance in a production? Compare your finding with the one from association rules? 
```{r}

```

